Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-15 bbbq_article.tex
Process feedback @bogeert. Thanks Geert! Progress #179

Our immune system fights microbial pathogens on a daily basis,
such as fungi, bacteria or viruses.
The acquired immune response
needs time to develop its specialized and more effective
combat forces and learn to recognize the pathogens 
by use of the Major Histocompatibility Complex (MHC).

@@

Antigen-presenting cells such as dendritic cells
use the MHC class I molecules to 
present randomly sampled peptides fragments,
as created by the immunoproteasome (\cite{chapiro2006destructive}).
If a cell gets infected by an intracellular pathogen, also the virus' peptides
will also be presented at the cell surface. The foreign viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

@@

All pathogens can be detected by the foreign proteins on 
their (bacterial or fungal) cell walls or (viral) envelope.
For bacteria and fungi, this is the main mode of their detection,
as these do not infect a cell with their (foreign) DNA.
T and B cells express MHC-II to detect foreign peptide fragments.
An immune response is started when MHC-II detects a pathogen.

@@

Any human's immune system detects only a fraction of all possible
peptide fragments.
For humans, the MHC proteins are encoded by the
HLA (Human Leukocyte Antigens) genes.
There are three genes encoding for MHC-I, which are HLA-A, HLA-B and HLA-C,
for MHC-II there are three major genes, which are HLA-DR, HLA-DQ, HLA-DP.
Each MHC complex can only bind a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
peptides they bind (\cite{lund2004definition}).
The HLA region of humans is highly polymorphic, with hundreds
to thousands of alleles for a gene (\cite{marsh2010nomenclature}).

@@

Because of these multiple and highly polymorphic MHC genes,
the number of pathogenic peptides that can be detected is increased,
as a wide variety of MHCs will be expressed,
each presenting their own subset of epitopes.

>>>>

Our immune system fights microbial pathogens, 
such as fungi, bacteria or viruses. 
An important part of the acquired immune response, 
that develops specialized and more effective recognition of pathogens, 
are T cells which recognize peptides derived from the pathogens 
by use of Major Histocompatibility Complexes (MHC).

@@

% GvdB: Not really relevant. You don’t really need to explain the role of DCs. What is far more important is the concept of what an MHC is: a protein complex that contains a fold, called the peptide-binding groove, which presents peptides. Also immediately explain that MHC-I presents peptides of about 9-amino acids in length to cytolytic T cells, whereas MHC-II presents longer peptides to helper T cells.

% Antigen-presenting cells such as dendritic cells
% use the MHC class I molecules to 
% present randomly sampled peptides fragments,
% as created by the immunoproteasome (\cite{chapiro2006destructive}).
% If a cell gets infected by an intracellular pathogen, also the virus' peptides
% will also be presented at the cell surface. The foreign viral antigen is detected 
% by cytotoxic T lymphocytes, which will kill the infected cell.

@@

% GvdB: Also not really relevant

% All pathogens can be detected by the foreign proteins on 
% their (bacterial or fungal) cell walls or (viral) envelope.
% For bacteria and fungi, this is the main mode of their detection,
% as these do not infect a cell with their (foreign) DNA.
% T and B cells express MHC-II to detect foreign peptide fragments.
% An immune response is started when MHC-II detects a pathogen.

@@

Any individual's immune system detects only a fraction of all possible
peptide fragments.
For humans, the MHC proteins are encoded by the
HLA (Human Leukocyte Antigens) genes.
There are three genes encoding for MHC-I, which are HLA-A, HLA-B and HLA-C,
for MHC-II there are three major genes, which are HLA-DR, HLA-DQ, HLA-DP.
Each MHC complex can only bind a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
peptides they bind (\cite{lund2004definition}).
The HLA region of humans is highly polymorphic, with hundreds 
to thousands of different alleles, 
and each different HLA gene is called 
an MHC haplotype (\cite{marsh2010nomenclature}).

@@

Because of these multiple and highly polymorphic MHC genes,
the number of pathogenic peptides that can be detected is increased,
as a wide variety of MHCs will be expressed,
each presenting their own subset of epitopes.
It is believed that this improves immunity at the population level, 
as mutations in a protein that disrupt a particular MHC presentation, 
so-called escape mutations, 
will not affect MHC presentation for all haplotypes \richel{Reference needed}.
